---
title: "fatal-crash-findings"
author: "Sahana Jayaraman"
date: "12-1-2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(widgetframe_self_contained = FALSE) 
knitr::opts_chunk$set(widgetframe_isolate_widgets = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
# Findings on fatal crashes 
This RMD serves as documentation of analytical findings for a data-driven story on D.C. traffic deaths. The analysis below shows that fatalities in the District have been steadily rising for more than a decade, despite the city's implementstion of a strategy to eliminate deaths by 2024. Lower-income communities east of the Anacostia river bore the brunt of the impact: Across the eight years analyzed, nearly half of all deaths occurred in Wards 7 and 8. 

Four of the five neighborhoods with the most deaths over the past eight years are home to majority-Black residents. Meanwhile, five majority-white and higher-income neighborhoods had no traffic deaths during the period analyzed. 

## load libraries
```{r}

library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(readxl)
library(writexl)
library(sf)
library(leaflet)
library(htmltools)
library(viridis)
library(DT)
library(vroom)

```

## Load and clean data
```{r}
# fatalities data: acquired from DDOT, with one added by OCME
fatalities_ddot <- read_xlsx("../../data/clean/v0/input/11_15_ccns_ddot.xlsx")

# census tracts, 2019 
census_tracts_2019 <-  st_read("../../data/clean/v0/input/tl_2019_11_tract/tl_2019_11_tract.shp") %>%
  clean_names()

# demographics from 2020 census, refit to 2010 census tracts by Ted  
demographics_2019_tracts <- read_csv("../../data/clean/v0/input/dc_2019_tracts_short.csv") %>%
  clean_names()

# tracts-to-hpn crosswalk: This  is a crosswalk matching 2010 D.C. census tracts to D.C. Health planning neighborhoods
tracts_to_neighborhoods <- read_csv("../../data/clean/v0/input/neighborhoods_to_tracts.csv") %>%
  clean_names() %>%
  mutate(geoid = as.character(geoid))

# shapefile of D.C. wards 
wards_shp <- st_read("../../data/clean/v0/input/Ward_from_2012/Ward_from_2012.shp") %>%
  clean_names() 

# shapefile of D.C. health planning neighborhoods
neighborhoods_shp <- st_read("../../data/clean/v0/input/DC_Health_Planning_Neighborhoods/DC_Health_Planning_Neighborhoods.shp") %>%
  clean_names() %>% 
  mutate(name = case_when(
    name == "N 24 FOGGY BOTTOM/GWU" ~ "N24 FOGGY BOTTOM/GWU",
    TRUE ~ name
  )) # the extra space here messes with analysis. Removing to make this uniform. 

# Setting crs of neighborhood and ward data to match the NAD83 data from every other file 
neighborhoods_shp <- st_transform(neighborhoods_shp, st_crs(4269))

wards_shp <- st_transform(wards_shp, st_crs(4269))
  

# medical examiners records of traffic fatalities and DDOT fatalities data, joined after creation of hand-checked crosswalk
me_ddot_joined <- vroom("../../data/clean/v0/output/me_ddot_joined.csv") 

crashes <- read_rds("../../data/clean/v0/input/df_crashes.rds") %>% 
  clean_names() %>%
  mutate(date = as_date(fromdate),
         year = year(date),
         month = month(date),
         day = day(date)) %>%
  rename(fatal_passenger = fatalpassenger) %>% 
  # distincting on ccn removes duplicate crashes; ccn is the report number associated with an individual crash. Checked crash numbers with DDOT and PIO confirmed. 
  distinct(ccn, .keep_all = TRUE) %>%
  filter(year >= "2014") 


```

## Clean fatalities and display as data table
```{r}

fatalities_ddot_clean <- fatalities_ddot %>%
  clean_names() %>%
  mutate(crash_date_clean = as_date(crash_date_and_time), 
         year = year(crash_date_clean)) %>%
  select(-c("death_case_number","jurisdiction", "death_case_year" )) %>%
  filter(year >= "2014") %>%
  mutate(ccn_clean = str_remove_all(ccn, "-"),
         id = row_number(), .before = ccn,  
         address_location = tolower(address_location)) 

# For reporters to sort and filter
datatable(fatalities_ddot_clean)

```

## Fatalities weighted by yearly crashes

Findings: 2020 and 2021 particularly interesting; they have the least crashes and the most deaths per crash
```{r}

# yearly total crashes, confirmed by PIO 
crashes_grouped <- crashes %>% 
  group_by(year) %>% 
  summarise(number_crashes = n())

# yearly fatalities 
fatalities_grouped <- fatalities_ddot_clean %>%
  group_by(year) %>%
  summarise(number_fatalities = n())

# fatalities by 5k crashes 
fatalities_by_5k_crash <- fatalities_grouped %>% 
  left_join(crashes_grouped, by = "year") %>%
  mutate(fatalities_per_5k_crashes = (number_fatalities/number_crashes)*5000) 
  
datatable(fatalities_by_5k_crash)

```
## Breakdown on the death types per year 

Findings: Pedestrians saw the most fatalities across the years 
```{r}

death_type <- fatalities_ddot_clean %>% 
  group_by(death_type, year) %>%
  mutate(death_type = case_when(death_type == "sco" ~ "scooter", TRUE ~ death_type)) %>%
  summarise(number_individuals = n()) %>%
  pivot_wider(names_from = "year", values_from = "number_individuals") %>%
  mutate_at(c(2:9), ~replace(., is.na(.), 0)) %>% 
  clean_names() %>% 
  rowwise() %>%
  mutate(total_deaths = sum(x2014 + x2015 + x2016 + x2017 + x2018 + x2019 + x2020 + x2021)) %>%
  arrange(desc(total_deaths))

datatable(death_type)

```

## Breakdown of the people who died 

Findings: We don't know the race of 46 of the people who died — but of the folks we do know, the majority of the dead are Black. At least 144 of the 250 people who died were Black; that's likely an undercount. 
```{r}
# by race 
death_race <- me_ddot_joined %>% 
  group_by(race) %>% 
  summarise(number_victims = n()) %>%
  arrange(desc(number_victims))

datatable(death_race)
#datatable(death_race_type)
```
## Deaths in age groups 

Findings: More than 40% of the people who died were between 20 and 40. 
```{r}
deaths_age_groups <- me_ddot_joined %>% 
  mutate(age_group =  case_when(age_clean>= 0 & age_clean < 10 ~ "<10",
                                age_clean>= 10 & age_clean<= 19 ~ "10-19",
                                age_clean>= 20 & age_clean<= 29 ~ "20-29",
                                age_clean>= 30 & age_clean<= 39 ~ "30-39",
                                age_clean>= 40 & age_clean<= 49 ~ "40-49",
                                age_clean>= 50 & age_clean<= 59 ~ "50-59",
                                age_clean>= 60 & age_clean<= 69 ~ "60-69",
                                age_clean>= 70 & age_clean<= 79 ~ "70-79",
                                age_clean>= 80 & age_clean<= 89 ~ "80-89",
                                age_clean>= 90 ~ "90+")) 

deaths_age_summary <- deaths_age_groups %>% group_by(age_group) %>% count() 

# breakdown by 5 years, to test 
deaths_age_groups_shorter <- me_ddot_joined %>% 
  mutate(age_group =  case_when(age_clean>= 0 & age_clean < 5 ~ "<5",
                                age_clean>= 5 & age_clean<= 9 ~ "5-9",
                                age_clean>= 10 & age_clean<= 14 ~ "10-14",
                                age_clean>= 15 & age_clean<= 19 ~ "15-19",
                                age_clean>= 20 & age_clean<= 24 ~ "20-24",
                                age_clean>= 25 & age_clean<= 29 ~ "25-29",
                                age_clean>= 30 & age_clean<= 34 ~ "30-34",
                                age_clean>= 35 & age_clean<= 39 ~ "35-39",
                                age_clean>= 40 & age_clean<= 44 ~ "40-44",
                                age_clean>= 45 & age_clean<= 49 ~ "45-49",
                                age_clean>= 50 & age_clean<= 54 ~ "50-54",
                                age_clean>= 55 & age_clean<= 59 ~ "55-59",
                                age_clean>= 60 & age_clean<= 64 ~ "60-64",
                                age_clean>= 65 & age_clean<= 69 ~ "65-69",
                                age_clean>= 70 & age_clean<= 74 ~ "70-74",
                                age_clean>= 75 & age_clean<= 79 ~ "75-79",
                                age_clean>= 80 & age_clean<= 84 ~ "80-84",
                                age_clean>= 85 & age_clean<= 89 ~ "85-89",
                                age_clean>= 90 & age_clean<= 94 ~ "90-94",
                                age_clean>= 95 ~ "95+")) %>%
  group_by(age_group) %>% 
  count() %>% 
  arrange (desc(n))

datatable(deaths_age_summary)
datatable(deaths_age_groups_shorter)
```

## Cluster map of each individual crash, across dc and by road

Findings: Roads identified with some of the most deaths — NY AVE. NE, SOUTHERN AVE SE. and ALABAMA AVE SE. 
```{r}

leaflet(fatalities_ddot_clean) %>%
  addTiles(options = providerTileOptions(minZoom = 10, maxZoom = 20)) %>% 
  addMarkers(lat = ~y, lng = ~x,
             label = ~id,
              clusterOptions = markerClusterOptions()) 

# create a df that filters for the word you're looking for in the data 
# Variable road is the name of the road you're looking for. Can be regex 

fatalities_road <- function(road){
df <- fatalities_ddot_clean %>%
  filter(str_detect(address_location, road))

leaflet(df) %>%
  addTiles(options = providerTileOptions(minZoom = 10, maxZoom = 20)) %>% 
  addMarkers(lat = ~y, lng = ~x,
             label = ~id,
              clusterOptions = markerClusterOptions()) 
}

fatalities_road("southern")

```

## Deaths by neighborhood 

Note: 10 of the deaths were **right** along the border of DC, just beyond the boundaries of the city's shapefile. Using a census tract map, I placed those fatalities into the census tracts they would be in (and thus, the neighborhoods they'd be in). MPD responded to them all, and they're all counted in DDOT's records of traffic deaths in DC. For those on the border of two tracts, I tested the analysis with the death counted in either tract to ensure it wouldn't affect findings. 
```{r}

crashes_shp <- st_as_sf(fatalities_ddot_clean, coords = c("x", "y"), crs = 4269)

deaths_with_geoid <- st_join(crashes_shp, census_tracts_2019) %>%
  mutate(is_border = ifelse(
    is.na(geoid), "y", "n"
  )) %>%
  mutate(geoid = ifelse(id == "1", "11001006202",
                        ifelse(id == "5", "11001009508",
                               ifelse(id == "30", "11001011100",
                                      ifelse(id == "118", "11001001001",
                                             ifelse(id == "206", "11001007304",
                                                    ifelse(id == "227", "11001009902",
                                                           ifelse(id == "231", "11001007603",
                                                                  ifelse(id == "239", "11001007409",
                                                                         ifelse(id == "242", "11001007707",
                                                                                ifelse(id == "246", "11001009811", geoid)
                                                                                ))))))))))%>%
  select(id, ccn, death_type, address_location, ward, crash_date_and_time, crash_date_clean, year, ccn_clean, geoid, is_border, geometry)

# join deaths with geoid to tracts-to-neighborhoods crosswalk to get deaths with census tract and neighborhood designations  
fatalities_neighborhoods <- deaths_with_geoid %>%
  left_join(tracts_to_neighborhoods, by = "geoid")

datatable(fatalities_neighborhoods)

# count number of deaths by neighborhood over all years
deaths_by_neighborhood <- fatalities_neighborhoods %>%
  group_by(hpn_label, code) %>%
  count() %>%
  arrange(desc(n))

# count number of deaths by neighborhood year-over-year 
death_by_neighborhood_year <- fatalities_neighborhoods%>%
  group_by(hpn_label, code, year) %>%
  count()

# transforming deaths by neighborhoods by year into a dataframe 
death_by_neighborhood_year <- as.data.frame(death_by_neighborhood_year)%>%
  select(-geometry)%>%
  pivot_wider(names_from = "year", values_from = "n") %>%
  mutate_if(is.numeric, as.character) %>%
  mutate_at(c(3:10), ~replace(., is.na(.), 0)) %>% 
    mutate_at(3:10, as.numeric) %>%
  pivot_longer(cols = 3:10, names_to = "year", values_to = "n" ) 

# year-over-year deaths by neighborhood and total deaths, organized for display 
dbny_display <- death_by_neighborhood_year %>% 
  pivot_wider(names_from = "year", values_from = "n")%>% 
  clean_names() %>% 
  relocate(x2019, .before = x2020) %>% 
  rowwise() %>%
  mutate(total_deaths = sum(x2014 + x2015 + x2016 + x2017 + x2018 + x2019 + x2020 + x2021))

# see deaths by neighborhood, by year
datatable(dbny_display)

```

## Death types by neighborhood 
```{r}

death_type_neighborhood <- as.data.frame(fatalities_neighborhoods) %>%
  group_by(death_type, hpn_label) %>%
  count() %>%
  select(1,2,3) 

datatable(death_type_neighborhood)

```

## Demographics by neighborhood
```{r}
# join census demographics to tracts-to-neighborhoods 
demographics_tracts <- tracts_to_neighborhoods%>%
  mutate(geoid = as.numeric(geoid)) %>%
  left_join(demographics_2019_tracts, by = "geoid")

# get populations by neighborhood
pops_by_nbh <- demographics_tracts %>%
  group_by(hpn_label, code) %>%
  select(-c("hu20", "occhu20")) %>%
  summarise_at(vars(7:26), sum) %>%
  select(1:6, 13) %>%
  mutate(percent_black = (black20/totpop20)*100) %>%
  mutate(percent_hisp = (hisp20/totpop20)*100) %>%
  mutate(percent_white = (white20/totpop20)*100) %>%
  mutate(percent_kid = ((totpop20 - a_totpop20)/totpop20)*100)


# join deaths  them both to neighborhoods_shp
neighborhoods_shp_joined <- neighborhoods_shp %>%
  left_join(as.data.frame(deaths_by_neighborhood), by = "code") %>%
  left_join(pops_by_nbh, by = "code") %>%
  mutate(n = as.character(n),
    n = case_when(
    is.na(n) ~ "0",
    TRUE ~ n
  ),
  n =  as.numeric(n))

```

## Fatalities by ward — a few different ways
Wards 7 and 8, together, have about 45% percent of all deaths. Ward 8 has more deaths than wards 6, 4 and 1 put together. 
```{r}

# grouping and counting deaths by wards as encoded by DDOT
# note: DDOT mis coded 2 deaths into the wrong wards. Ward 7 had 53 deaths, not 54
fatal_ward <- fatalities_ddot_clean %>% 
  group_by(ward) %>% 
  count() %>% 
  arrange(desc(n))

datatable(fatal_ward)

# checking DDOT ward encoding with a spatial join, points (fatalities) to polygons (wards)
deaths_by_ward_shp <- crashes_shp  %>%
  st_join(wards_shp) %>%
  group_by(ward.y) %>%
  count()

# Leaflet choropleth of deaths by wards, as visual aid
fatal_ward_polygon <- wards_shp %>% 
  st_join(deaths_by_ward_shp) %>% 
  rename("deaths" = "n")

pal <- colorNumeric("viridis", domain = fatal_ward_polygon$deaths, reverse = TRUE)

fatal_ward_polygon %>%
  leaflet() %>%
  addTiles(options = providerTileOptions(minZoom = 10, maxZoom = 20)) %>%
  addPolygons(fillColor = ~pal(deaths), color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.8, 
    highlightOptions = highlightOptions(color = "white", weight = 2,
    bringToFront = TRUE),
    label = ~ward,
    labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "8px",
    direction = "auto")) %>%
  addLegend(pal = pal, values = ~deaths, opacity = 0.8,
  position = "topright")


```

## Choropleth building

A function to build choropleths. 

**df**: a variable denoting the dataframe to use 

**column**: a variable denoting the column to use as the domain for the choropleth; numeric values that color shading should be determined by. Also used in hover labels

**label**: a variable denoting the column that should be used for the labels that appear when someone hovers over parts of the map

**title**: a variable purely for labeling, to denote what the figures in 'column' are measuring (deaths, demographics, etc.)
```{r}

heatmap <- function(df, column, label, title){

pal <- colorNumeric("viridis", domain = column, reverse = TRUE)

labels <- sprintf(
  "<strong>%s</strong><br/>%g %s",
  label, column, title
) %>% lapply(HTML)

df %>%
  leaflet() %>%
  addTiles(options = providerTileOptions(minZoom = 10, maxZoom = 20)) %>%
  addPolygons(fillColor = ~pal(column), color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.8, 
    highlightOptions = highlightOptions(color = "white", weight = 2,
    bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "8px",
    direction = "auto"))%>%
  addLegend(pal = pal, values = ~column, opacity = 0.8, title = title,
  position = "topright")
}
```

## Fatalities across neighborhoods and years

Findings: It appears more fatalities are ocurring in majority Black neighborhoods than in majority white ones. Neighborhoods with the most deaths were: Twining, the National Mall area, Saint Elizabeth's, Chinatown, Fort Licoln and Licoln heights. 
```{r}
heatmap(neighborhoods_shp_joined, neighborhoods_shp_joined$n, neighborhoods_shp_joined$dc_hpn_nam, "Fatalities" )

```

## Percent Black, per neighborhood
```{r}

heatmap(neighborhoods_shp_joined, neighborhoods_shp_joined$percent_black, neighborhoods_shp_joined$dc_hpn_nam, "percent Black")
```
## Percent white, per neighborhood
```{r}

heatmap(neighborhoods_shp_joined, neighborhoods_shp_joined$percent_white, neighborhoods_shp_joined$dc_hpn_nam, "percent white")

```

## Crashes, all years, individual maps

Findings: Year to year, the pattern of where crashes are ocurring changes a little bit. They remaiin concentrated in the same general area, however: Clustered around the commercial areas of U st., chinatown and the national mall and in residential neighborhoods that line the NE and SE sides.
```{r}

# year-by-year fatality + neighborhood + demographics data, for single-year and year-by-year maps
neighborhoods_shp_joined_yr <- neighborhoods_shp %>%
  left_join(death_by_neighborhood_year, by = "code") %>%
  left_join(pops_by_nbh, by = "code")

ggplot(neighborhoods_shp_joined_yr) +
  geom_sf(aes(fill = n), color = "black")+
  theme_void() +
  scale_fill_viridis(direction = -1) +
  facet_wrap(vars(year))

```

## Fatality maps by year 
```{r}

staticmap <- function(number){
  
neighborhoods_shp_joined_short <<- neighborhoods_shp_joined_yr%>%
  filter(year == number)
 
ggplot(neighborhoods_shp_joined_short) +
  geom_sf(aes(fill = n), color = "black")+
  theme_void() +
  scale_fill_viridis(direction = -1) 
} 
  
staticmap("2021")

```
